\chapter{Introduction}

\section{Background}
The background section will give a short introduction to the three main areas in this project; The technology that is used in vCare, the Danish Healthcare System and finally an introduction to the main target group who will be affected when introducing telemedicine in rehabilitation. 

\subsection{Technology}

The technology section will be based on the new ICT based virtual coaching solution vCare. 

The basic concept of vCare is carried out by a central eHealth platform that serves central infrastructure services. The platform obtains the information delivered by sensors or gained by the direct interaction between the patient and the virtual coach. The devices added to this platform consist of camera, microphones and a Kinect which makes the platform able to track movements. The information from the devices are conducted by a real-time processor. Beside the platform the infrastructure delivers supporting services to improve the quality of life of patients. The service provides physical and cognitive exercises as well as education material within nutrition and life behaviour. This service will be extended with a care pathway and a knowledge layer that enables personalized exercises and material for the given patient. Based on algorithms the virtual coach is flexible regarding the patients’ needs and hereby able to make specific rehabilitation programs. The platform can be implemented on different devices, e.g., tablets, smartphones, TV screens etc. \cite{Technical}. 


\subsection{The Danish Healthcare System}

The establishment of the Danish Healthcare System started in the eighteenth century. The first hospital was placed in Copenhagen and opened in 1757. This hospital is still running and today it is known as Rigshospitalet. Outside of the capital small hospitals were built during the late eighteenth century. At that time the hospitals were partly financed by taxes, patient payment and charity. In the late nineteenth century every thirteenth Dane was a member of a sick-benefit association which the Danish Government co-funded. The Danish Welfare State has roots in 1933 where the Social reform was founded. By this reform to all Danes with a low income it became a demand that they were members of a sick-benefit association. During the thirties taxes gradually became the dominant finance source at the Danish Healthcare System.\\ 
The sick-benefit associations were shut down in 1973 and replaced by public health insurance. The Danish public health insurance is paid by the Danes themselves by taxes. The insurance provides free care for everyone regardless of income and residence. This public health insurance includes hospital stays, surgery, visits to a GP and specialist'. Furthermore, it provides partly funding for dentist, physiotherapist, chiropractor, podiatrist and contributes to medicine.   \\

% Der mangler en kilde på dette afsnit

\subsection{Target Group and Market Segment}

By introducing telemedicine, the rehabilitation process is brought directly to patients' homes and mostly targets people with chronic conditions, which includes cardiac patients. Telemedicine rehabilitation is used to prevent hospitalization, to improve patients' feeling of safety, to empower patients to manage their own chronic condition and hereby improve patients' quality of life \cite{Emergence}. 

The need for cardiac rehabilitation is evaluated for all patients with heart disease. This includes both patients who have had a balloon dilation or by-pass surgery and patients with stable ischemic heart disease.
Patients with heart failure, pacemaker or patients who have had heart-valve surgery or cardiac transplantation are also being evaluated for the purpose of cardiac rehabilitation \cite{Rehabilitering}. By this statement it is seen, that this invention will involve a large target group.

To teach cardiac patients about their illness and how they are able to influence the course of the disease, results in a reduced risk of dying. Furthermore, research shows that rehabilitation programs with physical exercise reduce cardiac mortality \cite{Hjerteforening}.    
   

\section{Problem Statement}
More than half of the danish citizens over the age of 55 suffer from a cardiovascular disease. Furthermore, cardiovascular diseases are one of the most common causes to death in Denmark. The total cost of treating cardiovascular patients at the Danish Healthcare System was 5.5 billion DKK in 2015. Every year approximately 55.700 Danes is diagnosed with cardiovascular disease.   

Nearly 107.100 Danes are hospitalized every year for cardiovascular disease and almost 73.100 Danes are yearly at one or more consultations at the hospital. Approximately 23 percent of the cardiovascular patients are readmitted into the hospital within 30 days after being discharged. It has been proven, that cardiac rehabilitation results in a reduction in deaths caused by cardiovascular diseases and the need for readmissions \cite{Hjerteforening}.

All this indicates that cardiovascular patients constitute a large part of the Danish states economy. This leads to our problem statement which is:

\begin{itemize}
	\item What impact would an ICT solution for rehabilitation have on both cardiovascular patients and the Danish Healthcare System?
	\item How can ICT help to prevent readmission of cardiac patients?
	\item Which barriers/challenges can such system meet in implementation?
\end{itemize}

\subsection{Delimitation}
This project is limited only to be focusing on healthcare in Denmark and how the technology within rehabilitation will have an essential impact on the Danish Healthcare System. However, the project will be compared to related ICT solutions in EU as scientific articles based on The Danish Healthcare System is limited in this research area. 

Relevant data on how the Danish Healthcare System is establish will mainly be based on literature found in books and on websides where guidelines, statistics and the historical development is being published. 


